---
title: "Intro to Interactive Maps with Leaflet"
output:
  html_document:
    toc: true
    toc_float: true
    theme: "yeti"
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{css, echo=FALSE}
.footer {
    position: fixed;
    left: 0;
    right: 0;
    bottom: 0;
    width: 100%;
    text-align: right;
}


```	

<div class="footer">
  <p> mhughes@npr.org |  Updated  <script> document.write(new Date().toLocaleDateString()); </script> </p>
</div>
<style type="text/css">
h4.date {
  text-align: right;
}
</style>

---------------------------------------

# 1. Basics

---------------------------------------

##  Load packages and pull data




```{r load}
library(readr)
library(tidyverse)
library(leaflet)
cluster_demog<-read_csv(
"http://ec2-54-235-58-226.compute-1.amazonaws.com/storage/f/2014-02-23T17%3A11%3A51.737Z/dc-neighborhood-cluster-demographics.csv")
```



##  Blank Map


```{r blank}
leaflet( data = cluster_demog ) %>%
    addTiles() 
```




##  Add Markers
 Add markers with ``` addMarkers()```, specifying the columns for latitude, ```lat```, and longitude, ```lng```, at minimum.
```{r add_markers}
leaflet( data = cluster_demog ) %>%
    
    addTiles() %>%
    
    addMarkers(   lng = cluster_demog$lon_ctr      ### !!! ### 
                , lat = cluster_demog$lat_ctr        ### !!! ### 
                )                                  ### !!! ### 
```


##  Change Map "Tiles"

To do this, we swap the default ```addTiles()``` for  ``` addProviderTiles(providers$name_of_desired_tile).```  You can preview the built-in Leaflet tiles [here](http://leaflet-extras.github.io/leaflet-providers/preview/index.html).


```{r change_tiles}
leaflet( data = cluster_demog ) %>%
    
    addProviderTiles( providers$CartoDB.Positron ) %>%    ### !!! ### 
   
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat=cluster_demog$lat_ctr 
                )

```



```{r night_tiles}
leaflet( data = cluster_demog ) %>%
    
    addProviderTiles( providers$NASAGIBS.ViirsEarthAtNight2012 ) %>%    ### !!! ### 
    
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat = cluster_demog$lat_ctr 
                )

```

## Set the initial corrdinates and zoom level
The ```setView()``` utility function with three arguments: latitude, ```lat```, and longitude, ```lng```, and the zoom level, ```zoom```.  

```{r zoom_out}
 ### !!! ### 
leaflet( data = cluster_demog ) %>% 
    setView(lng = -77.0369, lat = 38.9072,  zoom=04) %>%           ### !!! ### 
    
    addProviderTiles( providers$NASAGIBS.ViirsEarthAtNight2012 ) %>%   
    
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat = cluster_demog$lat_ctr 
                )

```



---------------------------------------


# 2. Pop-ups

---------------------------------------

##  Add Pop-ups
To do this, we add ```popup = desired_popup_data_here``` as an argument into  ``` addMarkers()```.

```{r add_popups}
leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat=cluster_demog$lat_ctr 
                , popup =  cluster_demog$Full_Name       ### !!! ###     
                )  
                

```






##  Add More Info to Pop-ups
To do this, we just simply ```paste()``` the information together

```{r paste_popups}

 leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat=cluster_demog$lat_ctr 
                , popup =  paste(   cluster_demog$Full_Name                ### !!! ###     
                                  , cluster_demog$neighborhood_cluster ,sep=" "     ### !!! ### 
                                  )                                        ### !!! ### 
                )  
```







##  Format Pop-ups
Let's separate the pop-up details with a line break.   To do this, assign the ```sep``` argument of ```paste()``` to ``` sep = "<br/>"```

```{r format_popups}

 leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addMarkers(   lng = cluster_demog$lon_ctr 
                , lat=cluster_demog$lat_ctr 
                , popup =  paste(   sep = "<br/>"                     ### !!! ###     
                                  , cluster_demog$Full_Name          
                                  , cluster_demog$neighborhood_cluster 
                                  
                                  )
                )  
```

---------------------------------------

# 3. Markers

---------------------------------------

## From Plain Markers to Circle Markers
To do this, we will use ```addCircleMarkers()``` instead of ```addMarkers()```.  

```{r}
leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addCircleMarkers(   lng = cluster_demog$lon_ctr                      ### !!! ###
                , lat=cluster_demog$lat_ctr 
                , popup =  paste(   sep = "<br/>"                         
                                  , cluster_demog$Full_Name          
                                  , cluster_demog$neighborhood_cluster 
                                  
                                  )
                )  
```







## Circle Marker Radius


```{r}
 leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addCircleMarkers(   lng = cluster_demog$lon_ctr                      ### !!! ###
                , lat=cluster_demog$lat_ctr 
                , popup =  paste(   sep = "<br/>"                         
                                  , cluster_demog$Full_Name          
                                  , cluster_demog$neighborhood_cluster 
                                  )
                , radius = cluster_demog$`Population,_2010`
                )  

```

**Ew.** *Circles are so large that they fill the map!*



##   Circle Marker Radius


```{r}
 leaflet( data = cluster_demog ) %>%
    
    addProviderTiles(providers$CartoDB.Positron) %>%
    
    addCircleMarkers(   lng = cluster_demog$lon_ctr                     
                , lat=cluster_demog$lat_ctr 
                , popup =  paste(   sep = "<br/>"                         
                                  , cluster_demog$Full_Name          
                                  , cluster_demog$neighborhood_cluster 
                                  )
                , radius = cluster_demog$`Population,_2010`/5000              ### !!! ###
                )  

```







---------------------------------------

# 4. Choropleths, GeoJSON, Spatial Polygons

---------------------------------------

*Oh My!*

You might want this later: https://rstudio.github.io/leaflet/choropleths.html



## Using package ```geojsonio```


```{r}
#library(geojsonio)
cluster_outlines<-geojsonio::geojson_read("http://data.codefordc.org/dataset/8c0a1a57-8abf-4a29-abbb-ee2b890a878e/resource/3f9ce2d9-b438-4dab-a232-694cb80c7964/download/neighborhoodclusters.geojson", what = "sp")

cluster_outlines$Cluster_num<-as.numeric(grep("([0-9]+).*$",  as.character(cluster_outlines$NAME)))
```


```{r}
leaflet(cluster_outlines)  %>%
  addTiles() %>% 
    addPolygons()
```


## Color between the lines

Choosing to color by 2010 population.

```{r}
#summary(cluster_demog$`Population,_2010`)
cluster_outlines@data<-merge(cluster_outlines@data, cluster_demog, by="Cluster_num")

cut_offs<-data.frame(bins=quantile(cluster_outlines$`Population,_2010`, c(0,.2,.3,.4,.5,.6,.7,.8,1)))
bins <-c(0,cut_offs$bins)
pal <- colorBin("YlOrRd", domain = cluster_outlines$`Population,_2010`, bins = bins)



leaflet(cluster_outlines)  %>%
  addTiles() %>% 
    addPolygons(fillColor = ~pal(`Population,_2010`)                 ### !!! ###
     )


```

## Make it easier to see


```{r}
leaflet(cluster_outlines)  %>%
  addTiles() %>% 
   addPolygons(fillColor = ~pal(`Population,_2010`)  
      ,weight = 2,                                       ### !!! ###
  opacity = 1,                                            ### !!! ###
  color = "white",                                        ### !!! ###
#  dashArray = "3",
  fillOpacity = 0.7)

```

## Another option

```{r}
leaflet(cluster_outlines)  %>%
  addProviderTiles(providers$CartoDB.Positron) %>%              ### !!! ###
   addPolygons(fillColor = ~pal(`Population,_2010`)  
      ,weight = 1,
  opacity = 1,
  color = "white",
  
 # dashArray = "3",
  fillOpacity = 0.6)

```



## Hover Acation!

```{r}
leaflet(cluster_outlines)  %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
   addPolygons(fillColor = ~pal(`Population,_2010`)  
      ,weight = 1,
  opacity = 1,
  color = "white",
  fillOpacity = 0.6,
  highlight = highlightOptions(                     ### !!! ###
    weight = 2,                                     ### !!! ###
    fillOpacity = 0.75,                             ### !!! ###
    bringToFront = TRUE)                            ### !!! ###
  )

```




## Pop-ups

```{r}
labels <- sprintf(
  "<strong>%s</strong><br/>Population: %g",
  cluster_outlines$NBH_NAMES, cluster_outlines$`Population,_2010`
) %>% lapply(htmltools::HTML)



leaflet(cluster_outlines)  %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
   addPolygons(fillColor = ~pal(`Population,_2010`)  
      ,weight = 1,
  opacity = 1,
  color = "white",
  fillOpacity = 0.6,
  
  highlight = highlightOptions(
    weight = 2,
    fillOpacity = 0.75,
    bringToFront = TRUE)
  ,label = labels                                ### !!! ###
  )

```


## If the ```geojsonio``` package won't install

If you can't get ```geojsonio``` to work, you will want to try using the ```jsonlite``` package paired with the function ```readLines()```.  This example is incomplete, but will at least get you past the first step of reading in the geojson. 


```{r, eval=FALSE}

library(jsonlite)

geojson <- readLines("http://data.codefordc.org/dataset/8c0a1a57-8abf-4a29-abbb-ee2b890a878e/resource/3f9ce2d9-b438-4dab-a232-694cb80c7964/download/neighborhoodclusters.geojson"
    , warn = FALSE) %>%
  paste(collapse = "\n") %>%
  fromJSON(simplifyVector = FALSE)
  
```

---------------------------------------

# 5. More Resources

---------------------------------------


https://cran.r-project.org/web/packages/leaflet/leaflet.pdf

https://rstudio.github.io/leaflet/

https://rstudio.github.io/leaflet/morefeatures.html

Choropleths:

https://rstudio.github.io/leaflet/choropleths.html

Getting Lat-Longs

https://www.kaggle.com/joeleichter/us-zip-codes-with-lat-and-long


---------------------------------------








